---
title: "Class 13 GH"
author: "Emily R. Paris"
date: "2/18/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Download and clean HIV-Pr receptor structure

Here we wil download the 1HSG PDB structure from the online database, remove the bound drug from Merk and write out a protein only and lignd only new set of PDB files. **This code can be used on any protein, just change files!**

```{r}
library(bio3d)

# Download file to workspace
file.name <- get.pdb("1hsg")
```

```{r}
# Read file into R
hiv <-read.pdb(file.name)

# Call information about file
hiv
```

>Q1: What is the name of the two non protein resid values in this structure? What does resid
correspond to and how would you get a listing of all reside values in this structure? 
    
    The two non-protein resid values are called "HOH" and "MK1". This corresponds to parts of the small molecule (residuals) that are not proteins. Use 'atom.select'.

```{r}
# Select specifically for proteins
prot <- atom.select(hiv, string = "protein", value = TRUE)

# Make a file 
write.pdb(prot, file = "1hsg.protein.pdb")
```

```{r}
# Again for ligand
lig <- atom.select(hiv, string = "ligand", value = TRUE)
write.pdb(lig, file = "1hsg.ligand.pdb")
```


```{r}
# Sanity Check: make sure everything in 'lig' says "MK1" to make sure it all sorted to right spot
lig
```


