---
title: "Class 12: Structural Bioinformatics pt. 1"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## PDB Statistics

Here we inspect the types of structures in the main database for 3D biomolecular data - the PDB

>Q1: Determine the percentage of structures solved by X-Ray and Electron Microscopy. 

```{r}
# Read in file
stats <- read.csv("Data Export Summary.csv", row.names = 1)

# Calculate percentage values for each 'method'
ans <- stats$Total/sum(stats$Total) * 100
names(ans) <- rownames(stats)
round(ans, 2)
```


>Q2: Can you determine what proportion of structures are protein? Aim to have a rendered GitHub document with working code that yields your answers.

```{r}
round(sum(stats$Proteins)/sum(stats$Total), 2)
```

>Q2: Type HIV in the PDB website search box on the home page and determine how many
HIV-1 protease structures are in the current PDB?

#Part 2

Read a single PDB structure into R:
```{r}
library(bio3d)
pdb <- read.pdb("1hsg")
pdb
```

#Select Protein Only atoms with atom.select()

```{r}
prot <- atom.select(pdb, string = "protein", value = TRUE)
lig <- atom.select(pdb, string = "ligand", value = TRUE)
```

And write out these new PDB objects with the 'write.pdb()' function.

```{r}
#This is less error prone and more reproducible than copy and paste.
write.pdb(prot, file = "1hsg_protein.pdb")
write.pdb(lig, file = "1hsg_ligand.pdb")
```


```{r}
attributes(pdb)
```

```{r}
pdb$atom
```


>Q3: Water molecules normally have 3 atoms. Why do we see just one atom per water molecule
in this structure?

They're too small! Resolution = 2 Angstroms, but Hydrogen is smaller than that.

>Q4: There is a conserved water molecule in the binding site. Can you identify this water
molecule? What residue number does this water molecule have (see note below)?

















